Parallel Adaptive Importance Sampling

### 2016-12-04 PAIS.tex
changed algorithm style

\documentclass[final]{siamltex}
%test change
\usepackage{cite}
\usepackage{graphicx,bbm,pstricks}
\usepackage{pifont}
\usepackage{bbm,algorithmic,mdframed,placeins,multirow,booktabs,subfigure}
\usepackage[ruled,vlined,linesnumbered,algosection]{algorithm2e}

@@

The proposal distribution could be constructed in many different ways,
but we choose to use a mixture distribution, made up of a sum of MCMC
proposal distributions for each of the members of the ensemble. Once the proposal is
constructed, we can sample a new set of states from the proposal
distribution, and each is assigned a weight given by the ratio of the target
density and the proposal mixture distribution density. Assuming that
our proposal distribution is a good one, then the variance of the
weights will be small, and we will have many useful samples. Finally, we
need to create a set of evenly weighted samples which best represent
this set of weighted samples. This is achieved by implementing a
resampling algorithm. Initially we will use the
ETPF algorithm\cite{reich2013nonparametric}, although we will suggest an alternative strategy in
Section \ref{sec:AMR}. The output of the resampling algorithm gives us
a set of evenly weighted
samples that we believe represents the target distribution well, and
from this point we
can iterate the process once again. The algorithm is summarised in Table \ref{tab:PAIS}.

@@

\begin{table}
\begin{mdframed}
\begin{algorithmic}
\STATE $\X^{(0)} = \X_0 = [x_1^{(0)},x_2^{(0)},\ldots,x_M^{(0)}]^T$
\FOR{$i=0,1,2, ..., N-1$}
\STATE $\Y^{(i)} = [y_1^{(i)},y_2^{(i)},\ldots,y_M^{(i)}]^T, \quad y_j^{(i)} \sim
\nu(\cdot;x_j^{(i)})$
\STATE $\chi(y;\X^{(i)}) = \frac{1}{M}
\sum_{j=1}^M \nu(y;x_j^{(i)})$.
\STATE $\W^{(i)} = [w_1^{(i)},w_2^{(i)},\ldots,w_M^{(i)}]^T,$ \quad $w^{(i)}_j =
\frac{\pi(y_j^{(i)})}{\chi(y_j^{(i)};\X^{(i)})}$.
\STATE Resample: $(\W^{(i)},\Y^{(i)}) \rightarrow (\frac{1}{M}\mathbf{1}, \X^{(i+1)})$
\ENDFOR 
\end{algorithmic}
\end{mdframed}

\caption{A pseudo-code representation of the Parallel Adaptive
  Importance Sampler (PAIS).}
\label{tab:PAIS}
\end{table}

@@

\begin{table}[htb]
\begin{mdframed}
\begin{algorithmic}
\STATE $\b{z} = M\bar{\b{w}}$
\FOR{$i=1,2,\ldots,M$}
\STATE $J = \argmax_j z_j$
\STATE $p_{i,J} = \min\{1,z_J\}$
\STATE $z_J = z_J - p_{i,J}$
\WHILE{$\sum_j p_{i,j} <1$}
\STATE $K = \argmin_{k \in \{k|z_k>0\}} \|y_J - y_k\|$
\STATE $p_{i,K} = \min\{1-\sum_j p_{i,j}, z_K\}$
\STATE $z_K = z_K - p_{i,K}$
\ENDWHILE
\STATE $x_i = \sum_k p_{i,k}y_k$.
\ENDFOR
\end{algorithmic}
\end{mdframed}
\caption{The approximate multinomial resampler (AMR) algorithm.\label{tab:AMR}}
\end{table}

Table \ref{tab:AMR} describes the basis of the algorithm with
deterministic resampling, using the means of each of the
sub-multinomials as the new samples. This resampler was designed with the aims of being
numerically cheaper than the ETPF, and more accurate than straight multinomial
resampling. Therefore we now present numerical examples which
demonstrate this.

>>>>

\documentclass[final]{siamltex}
%test change
\usepackage{cite}
\usepackage{graphicx,bbm,pstricks}
\usepackage{pifont}
\usepackage{bbm,algorithmic,mdframed,placeins,multirow,booktabs,subfigure}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}

@@

The proposal distribution could be constructed in many different ways,
but we choose to use a mixture distribution, made up of a sum of MCMC
proposal distributions for each of the members of the ensemble. Once the proposal is
constructed, we can sample a new set of states from the proposal
distribution, and each is assigned a weight given by the ratio of the target
density and the proposal mixture distribution density. Assuming that
our proposal distribution is a good one, then the variance of the
weights will be small, and we will have many useful samples. Finally, we
need to create a set of evenly weighted samples which best represent
this set of weighted samples. This is achieved by implementing a
resampling algorithm. Initially we will use the
ETPF algorithm\cite{reich2013nonparametric}, although we will suggest an alternative strategy in
Section \ref{sec:AMR}. The output of the resampling algorithm gives us
a set of evenly weighted
samples that we believe represents the target distribution well, and
from this point we
can iterate the process once again. The algorithm is summarised in Algorithm \ref{alg:PAIS}.

@@

%\begin{table}
%\begin{mdframed}
%\begin{algorithmic}
%\STATE $\X^{(0)} = \X_0 = [x_1^{(0)},x_2^{(0)},\ldots,x_M^{(0)}]^T$
%\FOR{$i=0,1,2, ..., N-1$}
%\STATE $\Y^{(i)} = [y_1^{(i)},y_2^{(i)},\ldots,y_M^{(i)}]^T, \quad y_j^{(i)} \sim
%\nu(\cdot;x_j^{(i)})$
%\STATE $\chi(y;\X^{(i)}) = \frac{1}{M}
%\sum_{j=1}^M \nu(y;x_j^{(i)})$.
%\STATE $\W^{(i)} = [w_1^{(i)},w_2^{(i)},\ldots,w_M^{(i)}]^T,$ \quad $w^{(i)}_j =
%\frac{\pi(y_j^{(i)})}{\chi(y_j^{(i)};\X^{(i)})}$.
%\STATE Resample: $(\W^{(i)},\Y^{(i)}) \rightarrow (\frac{1}{M}\mathbf{1}, \X^{(i+1)})$
%\ENDFOR 
%\end{algorithmic}
%\end{mdframed}
%\caption{A pseudo-code representation of the Parallel Adaptive
%  Importance Sampler (PAIS).}
%\label{tab:PAIS}
%\end{table}

\begin{table}[!h]
\centering
\begin{algorithm}[H]
\DontPrintSemicolon
\BlankLine
	Set $\X^{(0)} = \X_0 = [x_1^{(0)},x_2^{(0)},\ldots,x_M^{(0)}]^\top$.\;
	\For {$i = 1, \dots, N$}
	{
		Sample $\Y^{(i)} = [y_1^{(i)},y_2^{(i)},\ldots,y_M^{(i)}]^\top, \quad y_j^{(i)} \sim
\nu(\cdot;x_j^{(i-1)})$.\;
		Calculate $\W^{(i)} = [w_1^{(i)},w_2^{(i)},\ldots,w_M^{(i)}]^\top,$ \quad $w^{(i)}_j =
\frac{\pi(y_j^{(i)})}{\chi(y_j^{(i)};\X^{(i-1)})}$, where
		\[
			\chi(y;\X^{(i-1)}) = \frac{1}{M}\sum_{j=1}^M \nu(y;x_j^{(i-1)}).
		\]

		Resample: $(\W^{(i)},\Y^{(i)}) \rightarrow (\frac{1}{M}\mathbf{1}, \X^{(i)})$.
	}
\caption{The parallel adaptive importance sampler (PAIS).\label{alg:PAIS}}
\end{algorithm}
\end{table}

@@

\begin{table}[!h]
\centering
\begin{algorithm}[H]
\DontPrintSemicolon
\BlankLine
	$\b{z} = M\bar{\b{w}}$.\;
	\For {$i = 1,\dots, M$}
	{
		$J = \argmax_j z_j$.\;
		$p_{i,J} = \min\{1,z_J\}$.\;
		$z_J = z_J - p_{i,J}$.\;
		\While {$\sum_j p_{i,j} <1$}
		{
			$K = \argmin_{k \in \{k|z_k>0\}} \|y_J - y_k\|$.\;
			$p_{i,K} = \min\{1-\sum_j p_{i,j}, z_K\}$.\;
			$z_K = z_K - p_{i,K}$.\;
		}
		$x_i = \sum_k p_{i,k}y_k$.\;
	}
\caption{The approximate multinomial resampler (AMR).\label{alg:AMR}}
\end{algorithm}
\end{table}

Algorithm~\ref{alg:AMR} describes the basis of the algorithm with
deterministic resampling, using the means of each of the
sub-multinomials as the new samples. This resampler was designed with the aims of being
numerically cheaper than the ETPF, and more accurate than straight multinomial
resampling. Therefore we now present numerical examples which
demonstrate this.
